---
title: "Tidy Tuesday 2021-26 Parks"
output:
  github_document: default
  html_notebook: default
  html_document:
    df_print: paged
    self_contained: false
bibliography: ["../tidytuesday.bib"]
nocite: '@*'
---

## Setup
```{r setup, message = FALSE, warning = FALSE, include = TRUE}
library(tidyverse)
```

## Load the data
```{r load_data}
# Data source: https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-22/parks.csv
parks_raw <- read_csv('data/parks.csv',
                col_types = cols(
                  year = col_factor(levels = NULL), # could cast as type date
                  rank = col_number(),
                  city = col_factor(levels = NULL), # for ease of processing
                  med_park_size_data = col_double(),
                  med_park_size_points = col_double(),
                  park_pct_city_data = col_character(), # trailing % sign
                  park_pct_city_points = col_double(),
                  pct_near_park_data = col_character(), # trailing % sign
                  pct_near_park_points = col_double(),
                  spend_per_resident_data = col_character(), # preceding $ sign
                  spend_per_resident_points = col_double(),
                  basketball_data = col_double(),
                  basketball_points = col_double(),
                  dogpark_data = col_double(),
                  dogpark_points = col_double(),
                  playground_data = col_double(),
                  playground_points = col_double(),
                  rec_sr_data = col_double(),
                  rec_sr_points = col_double(),
                  restroom_data = col_double(),
                  restroom_points = col_double(),
                  splashground_data = col_double(),
                  splashground_points = col_double(),
                  amenities_points = col_double(),
                  total_points = col_double(),
                  total_pct = col_double(),
                  city_dup = col_character(),
                  park_benches = col_double()
                )
              )
```
Question 1: Is it better to cast the year variable as a date, to be able to do arithmetic with it later?

## Explore and clean the data
### Remove stray characters and cast to appropriate data types
```{r clean_data}
parks <- parks_raw %>%
  mutate(
    park_pct_city_data = str_remove(park_pct_city_data, "%"), # remove %
    pct_near_park_data = str_remove(pct_near_park_data, "%"), # remove %
    spend_per_resident_data = str_remove(spend_per_resident_data, "\\$"), # remove $
    ) %>%
  mutate(
    park_pct_city_data = as.double(park_pct_city_data),
    pct_near_park_data = as.double(pct_near_park_data),
    spend_per_resident_data = as.double(spend_per_resident_data)
  ) %>%
  select(year:splashground_points, last_col(), everything(), -city_dup) # re-order cleanly
```
Question 2: Is there a better way to do these mutate calls more cleanly?

### Explore spread of values using total_points (total points per year)
```{r}
parks %>%
  select(year, total_points) %>%
  group_by(year) %>%
  ggplot(aes(x = year, y = total_points)) +
    geom_boxplot()
```
Seems the basis for calculating the index moves around significantly from year to year. Maybe the total percentage data looks better? Quite a few rows not used ...

### Explore spread of values using total_pct (percentage total per year)
```{r}
parks %>%
  select(year, total_pct) %>%
  group_by(year) %>%
  ggplot(aes(x = year, y = total_pct)) +
    geom_boxplot()
```

A bit more consistent, but let's see if we can fill that gap in 2017 ...

### Fill in gap in 2017 total_pct observations
2017 seems to be missing total percentages. Consulting the published datasheet (https://parkserve.tpl.org/mapping/historic/2017_ParkScoreRank.pdf) confirms that in 2017, the total_points were out of 100, so the total_point score is the same number as a percentage score. Let's fix that.

```{r 2017_gap}
parks <- parks %>%
  mutate(total_pct = case_when(
    (is.na(total_pct) == TRUE & year == "2017") ~ total_points, # points out of 100
    TRUE ~ total_pct)
  )
```

Let's have another look at that box plot of the total percentages

### Explore spread of values using total_pct (percentage total per year) - again
```{r}
parks %>%
  select(year, total_pct) %>%
  group_by(year) %>%
  ggplot(aes(x = year, y = total_pct)) +
    geom_boxplot()
```

That's much better. Let's go back and see about that total_points figure ...

### Explore spread of values using total_points (total points per year) - again
```{r}
parks %>%
  select(year, total_points) %>%
  group_by(year) %>%
  ggplot(aes(x = year, y = total_points)) +
    geom_boxplot()
```

### Re-calculate 2015 total_points observations
2015 seems to have very, very low average total_points. On closer inspection, it looks like the average points out of 20 for amentiies have been erroneously coded as total_points, rather than as one of several inputs summed to reach the total. Let's fix that.

```{r 2015_gap}
parks <- parks %>%
  mutate(total_points = case_when(
      year == "2015" ~ med_park_size_points + park_pct_city_points + pct_near_park_points + spend_per_resident_points + total_points, # points out of 120
      TRUE ~ total_points)
  )
```

### Fill in gaps in 2014 and 2019 total_points observations
And, 2019 and 2014 seem to be missing total_point observations. Drawing on the published datasheets for these years, it is possible to calculate total_points from the components and weightings. Let's rebuild these in the dataset.

```{r 2014_2019_gap}
parks <- parks %>%
  mutate(total_points = case_when(
      (is.na(total_points) == TRUE & year == "2014") ~ med_park_size_points + park_pct_city_points + pct_near_park_points + spend_per_resident_points + playground_points, # points out of 120
      (is.na(total_points) == TRUE & year == "2019") ~ med_park_size_points + park_pct_city_points + pct_near_park_points + spend_per_resident_points + amenities_points, # points out of 400
      TRUE ~ total_points)
  )
```

### Explore spread of values using total_points (total points per year) - again
```{r}
parks %>%
  select(year, total_points) %>%
  group_by(year) %>%
  ggplot(aes(x = year, y = total_points)) +
    geom_boxplot()
```

Much better. Looks like some changes to the methodology in 2018 and 2019. 

I wonder What happened to scores in 2017 ..?

Let's make all this a bit easier to re-produce in future ...

### Add in an index_base variable to help with future calculations
```{r add_index_base}
parks <- parks %>%
  mutate(index_base = case_when(
    year %in% c("2017") ~ 100,
    year %in% c("2012", "2013", "2014", "2015", "2016") ~ 120,
    year %in% c("2018") ~ 160,
    year %in% c("2019", "2020") ~ 400,
    )
  ) %>%
  select(year:total_points, last_col(), everything())
```



```{r explore_data, include = FALSE}
# Cruft ...
parks %>%
  select(year, city, total_points) %>%
  group_by(year) %>%
  summarize(n = n(), avg_score = mean(total_points, na.rm = TRUE))

# Looks like there are holes in 2019 and 2014 total points data ...
parks %>%
  select(year, total_points) %>%
  filter(year %in% c("2019", "2014"))

# Top 5 cities for each year, based on total percentage - which is more reliable than total points. 
parks %>%
  group_by(year) %>%
  select(year, rank, city, total_points, total_pct) %>%
  top_n(total_pct, n = 5)

# Number of times each city included
parks %>%
  group_by(city) %>%
  count(sort = TRUE)

# Number of unique cities
parks %>%
  distinct(city) %>%
  count()

# What's going on in 2015 with the total_count?

parks %>%
  select(year, ends_with("_points")) %>%
  filter(year == "2015")

```

```{r clumsy_points_over_time}
parks %>%
  select(year, city, ends_with("_points"), total_pct) %>%
  group_by(year) %>%
  mutate(year = as.Date(year, format = "%Y")) %>%
  summarize(avg_park_size = mean(med_park_size_points), avg_city_is_park = mean(park_pct_city_points), avg_park_access = mean(pct_near_park_points), avg_spend_per_res = mean(spend_per_resident_points), avg_amenities = mean(amenities_points), avg_total = mean(total_pct)) %>%
  ggplot(aes(year)) +
    geom_line(aes(y = avg_total, color = "total")) +
    geom_line(aes(y = avg_park_size, color = "size")) +
    geom_line(aes(y = avg_city_is_park, color = "% city is park")) +
    geom_line(aes(y = avg_park_access, color = "park access")) +
    geom_line(aes(y = avg_spend_per_res, color = "spend")) +
    geom_line(aes(y = avg_amenities, color = "amenities"))
```
```{r better_points_over_time}
parks %>%
  select(year, city, ends_with("_points"), total_pct) %>%
  group_by(year) %>%
  summarize(avg_park_size = mean(med_park_size_points), avg_city_is_park = mean(park_pct_city_points), avg_park_access = mean(pct_near_park_points), avg_spend_per_res = mean(spend_per_resident_points), avg_amenities = mean(amenities_points), avg_total = mean(total_pct)) %>%
  gather(key = "avg_metric", value = "value", avg_park_size:avg_total) %>%
  ggplot(aes(x = year, y = value, group = avg_metric)) +
    geom_line(aes(color = avg_metric))
```
## Explore where biggest changes in points over time

```{r}
rank_over_time <- parks %>%
  select(city, year, rank) %>%
  pivot_wider(names_from = year, values_from = rank) %>%
  mutate(
    `2019to2020` = replace_na(`2020` - `2019`, 0),
    `2018to2019` = replace_na(`2019` - `2018`, 0),
    `2017to2018` = replace_na(`2018` - `2017`, 0),
    `2016to2017` = replace_na(`2017` - `2016`, 0),
    `2015to2016` = replace_na(`2016` - `2015`, 0),
    `2014to2015` = replace_na(`2015` - `2014`, 0),
    `2013to2014` = replace_na(`2014` - `2013`, 0),
    `2012to2013` = replace_na(`2013` - `2012`, 0),
    total_variation = 
           abs(`2019to2020`) + 
           abs(`2018to2019`) +
           abs(`2017to2018`) +
           abs(`2016to2017`) +
           abs(`2015to2016`) +
           abs(`2014to2015`) +
           abs(`2013to2014`) +
           abs(`2012to2013`), 
    net_variation_2012to2020 = `2019to2020` + `2018to2019` +
                    `2017to2018` + `2016to2017` +
                    `2015to2016` + `2014to2015` +
                    `2013to2014` + `2012to2013`) %>%
  arrange(desc(total_variation))

rank_over_time %>%
  ggplot(aes(x = city, y = total_variation)) +
    geom_col() +
    coord_flip()

rank_over_time %>%
  top_n(10, wt = total_variation) %>%
  gather(key = "years", value = "rank_change", `2019to2020`:`2012to2013`) %>%
  select(city, years, rank_change, total_variation) %>%
  ggplot(aes(x = years, y = rank_change, group = city)) +
    geom_point() +
    geom_line(aes(color = city))

rank_over_time %>%
  top_n(-10, wt = total_variation) %>%
  gather(key = "years", value = "rank_change", `2019to2020`:`2012to2013`) %>%
  select(city, years, rank_change, total_variation) %>%
  ggplot(aes(x = years, y = rank_change, group = city)) +
    geom_point() +
    geom_line(aes(color = city))

rank_over_time %>%
  top_n(10, net_variation_2012to2020) %>%
  arrange(desc(net_variation_2012to2020)) %>%
  select(city, net_variation_2012to2020)
  
rank_over_time %>%
  top_n(10, net_variation_2012to2020) %>%
  gather(key = "years", value = "rank_change", `2019to2020`:`2012to2013`) %>%
  select(city, years, rank_change, total_variation, net_variation_2012to2020) %>%
  arrange(desc(net_variation_2012to2020)) %>%
  ggplot(aes(x = years, y = rank_change, group = city)) +
    geom_point() +
    geom_line(aes(color = city))


rank_over_time %>%
  top_n(-10, net_variation_2012to2020) %>%
  gather(key = "years", value = "rank_change", `2019to2020`:`2012to2013`) %>%
  select(city, years, rank_change, total_variation, net_variation_2012to2020) %>%
  arrange(desc(net_variation_2012to2020)) %>%
  ggplot(aes(x = years, y = rank_change, group = city)) +
    geom_point() +
    geom_line(aes(color = city))


```

## Test git commit


# References