---
title: "Tidy Tuesday "
output:
  html_document:
    df_print: paged
    self_contained: false
  github_document: default
  html_notebook: default
bibliography: ["../tidytuesday.bib"]
nocite: '@*'
---

# 

```{r setup, echo = FALSE}
library(tidytuesdayR)
library(tidyverse)
tuesdata <- tidytuesdayR::tt_load('2021-10-19')
pumpkins <- tuesdata$pumpkins
theme_set(theme_minimal())
```

# Data cleaning

```{r cleaning}
pumpkins_clean <- pumpkins %>% 
  filter(!grepl('damaged)', place)) %>% # remove dirty, broken rows
  separate(id, into = c("year", "type"), sep = "-") %>% # Split out id into more useful variables 
  mutate(weight_lbs = as.double(weight_lbs), weight_kgs = weight_lbs * 0.45359237, type = as.factor(type), place = as.numeric(place)) %>% 
  filter(!is.na(weight_lbs == TRUE) & !is.na(place == TRUE))
```


# EDA
## How many pumpkins per country>
```{r}
pumpkins_clean %>% 
  count(country, sort = TRUE)
```
So, looks like there are a whole lot of damaged entries in the data. We're talking about less than 50 rows, out of 28,065, so let's filter them out.

```{r}
pumpkins_clean %>% 
  filter(!grepl('damaged)', place))
```

## How many pumpkins of each type, average weight and range of weights?

```{r}
pumpkins_clean %>% 
  filter(type %in% c("F", "P", "S", "W", "L")) %>% 
  group_by(type) %>% 
  summarise(total = n(), avg_weight = mean(weight_lbs), min = min(weight_lbs), max = max(weight_lbs)) 
```

## What do the pumpkins looks like over time?

```{r}
pumpkins_clean %>% 
  group_by(year) %>% 
  summarise(total = n(), avg_weight = mean(weight_lbs), min = min(weight_lbs), max = max(weight_lbs)) 
```

# Plot the distribution of pumpkin weights, by type, over years

```{r}
pumpkins_clean %>% 
  filter(type != "T") %>% 
  group_by(year, type) %>% 
  select(year, type, weight_kgs) %>% 
  ggplot(aes(x = year, y = weight_kgs)) +
    geom_boxplot() +
    facet_wrap(~ type)

```
# Concentrate this view on the largest three of two types per year

```{r}
pumpkins_clean %>% 
  filter(place %in% c(1:10)) %>% 
  group_by(year, type) %>% 
  select(year, type, weight_kgs) %>% 
  ggplot(aes(x = year, y = weight_kgs)) +
    geom_boxplot() +
    facet_wrap(~ type)
```


## Find the grower with the most entries, and visualise over time

```{r}
pumpkins_clean %>% 
#  count(grower_name, sort = TRUE)
  filter(grower_name == "Risi, Robert") %>% 
  group_by(year, type) %>% 
  ggplot(aes(x = year, y = weight_kgs)) +
    geom_point() +
    facet_wrap(~ type)
```

# What about a network graph of the seed parents ..?

```{r}
pumpkins %>% 
  filter(!grepl('damaged)', place)) %>% # remove dirty, broken rows
  separate(id, into = c("year", "type"), sep = "-") %>%
  count(seed_mother, sort = TRUE)
```


```{r}
network_pumpkins <- pumpkins %>% 
  filter(!grepl('damaged)', place)) %>% # remove dirty, broken rows
  separate(id, into = c("year", "type"), sep = "-") %>%
  mutate(seed_mother = recode(seed_mother, 
                              Unknown = "unknown",
                              unknow = "unknown", 
                              'N/A' = "unknown",
                              'n/a' = "unknown",
                              Unkown = "unknown",
                              Unknow = "unknown",
                              UNKNOWN = "unknown",
                              '?' = "unknown",
                              Open = "open"),
        weight_lbs = as.double(weight_lbs),
        weight_kgs = weight_lbs * 0.45359237,              
    ) %>% 
  mutate(seed_mother = na_if(seed_mother, "unknown")) %>% 
  drop_na(seed_mother) %>%
  mutate(seed_mother = str_replace(seed_mother, "^\\S* ", "")) %>% 
  filter(!is.na(weight_lbs == TRUE)) %>% 
  group_by(type, year) %>% 
  summarize(total = n(), avg_weight_kg = mean(weight_kgs))
  #count(seed_mother, sort = TRUE)
  
```



# References

```{r clean_up}
rm(tuesdata, pumpkins)

```

