---
title: "Tidy Tuesday 2021-21 Ask a manager survey"
output:
  github_document: default
  html_notebook: default
  html_document:
    df_print: paged
    self_contained: false
bibliography: ["../tidytuesday.bib"]
nocite: '@*'
---

## Introduction


## Setup
```{r setup, message = FALSE}
library(tidytuesdayR)
library(tidyverse)
```

<!-- ## Import data - first time -->
<!-- ```{r get_the_data, cache = TRUE, message = FALSE} -->
<!-- survey <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-18/survey.csv') -->
<!-- write_csv(survey, 'data/survey.csv') -->
<!-- rm tuesdata -->
<!-- ``` -->

## Import data - subsequent times
```{r get_the_data_local, cache = TRUE, message = FALSE}
survey <- read_csv('data/survey.csv')
survey_clean <- survey
glimpse(survey)
```

## Data cleaning - this may take some time ...
Let's start by looking at the gender variable in our dataset.
```{r}
survey %>% 
  count(gender)
```
While this looks pretty clean, there are two immediate issues.
First, there is a single "prefer not to answer" observation even though there are numerous observations described as  "Other or prefer not to answer". Since there is no way to distinguish between survey respondents who are identifying with an "other" gender not listed and those that "prefer not to answer", this group of observations cannot be separated. To simplify, let's roll the single "prefer not to answer" observation into this broader group.
Second, these observations currently use the character data type, and it would be helpful for future analyses to have gender as a categorical variable.
Once tidied up, added a percentage column in case of future plots.
```{r clean gender}
survey_clean <- survey %>%
  mutate(gender_clean = fct_collapse(gender, "Other or prefer not to answer" = "Prefer not to answer")) %>%
  # clean up the temporary columns so variable is 'gender' and is of type categorical  
  select(-gender) %>%
  mutate(gender = as.factor(gender_clean)) %>%
  select(-gender_clean)
  
survey_clean %>%
  count(gender) %>%
  mutate (percent = round(n / sum(n) * 100, 1)) %>%
  arrange(desc(percent)) %>%
  ggplot(aes(x = reorder(gender, -percent), y = percent)) + geom_col()
```




I had cleaned this as strings, but it was pretty painful. Easier with forcats.


```




# Step 1: Fix up categories for age, education
# Step 2: Normalise/standardise currencies, create total salary
# Step 3: Clean gender into categories



## References
