---
title: "Tidy Tuesday 2021-21 Ask a manager survey"
output:
  github_document: default
  html_notebook: default
  html_document:
    df_print: paged
    self_contained: false
bibliography: ["../tidytuesday.bib"]
nocite: '@*'
---

## Introduction


## Setup
```{r setup, message = FALSE}
library(tidytuesdayR)
library(tidyverse)
library(countrycode)
```

<!-- ## Import data - first time -->
<!-- ```{r get_the_data, cache = TRUE, message = FALSE} -->
<!-- survey <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-18/survey.csv') -->
<!-- write_csv(survey, 'data/survey.csv') -->
<!-- rm tuesdata -->
<!-- ``` -->

## Import data - subsequent times
```{r get_the_data_local, cache = TRUE, message = FALSE}
survey <- read_csv('data/survey.csv')
survey_clean <- survey
glimpse(survey)
```

## Data cleaning - this may take some time ...
Let's start by looking at the gender variable in our dataset.
```{r}
survey %>% 
  count(gender)
```
While this looks pretty clean, there are two immediate issues.
First, there is a single "prefer not to answer" observation even though there are numerous observations described as  "Other or prefer not to answer". Since there is no way to distinguish between survey respondents who are identifying with an "other" gender not listed and those that "prefer not to answer", this group of observations cannot be separated. To simplify, let's roll the single "prefer not to answer" observation into this broader group.
Second, these observations currently use the character data type, and it would be helpful for future analyses to have gender as a categorical variable.
Once tidied up, added a percentage column in case of future plots.
```{r clean gender}
# Clean the gender data, by collapsing one category into another.
survey_clean <- survey %>%
  mutate(gender_clean = fct_collapse(gender, "Other or prefer not to answer" = "Prefer not to answer")) %>%
# Clean up the temporary columns so variable name is 'gender' and set data type to categorical.
  select(-gender) %>%
  mutate(gender = as.factor(gender_clean)) %>%
  select(-gender_clean)
  
# Calculate the percentages of each category, sort the data and plot it.
survey_clean %>%
  count(gender) %>%
  mutate (percent = round(n / sum(n) * 100, 1)) %>%
  arrange(desc(percent)) %>%
  ggplot(aes(x = reorder(gender, -percent), y = percent)) + geom_col()
```
I had cleaned this as strings, but it was pretty painful. Easier with forcats. To simplify the charty, it would be nice to be able to collapse the Other or prefer not to answer into NA ... another time.

Let's look at country now.

```{r}
survey %>% 
  count(country) %>%
  arrange(desc(n))
```
Let's have a go at tidying some of these up ...
```{r}

survey$country <- survey$country %>%
  str_remove_all("'.'") %>% # remove full stops
  str_to_lower()        # remove capitalisation inconsistencies

survey_clean <- survey %>%
  # select(country, state, city) %>% # Testing only, shows a bit more context for cleaning ...
  mutate(country = case_when(
    country %in% c("england", "englang", "scotland", "wales", "northern ireland", "united kindom", "unites kingdom", "uk", "u.k.") ~ "united kingdom",
    country %in% c("can", "canda", "csnada", "canad") ~ "canada",
    country == "danmark" ~ "denmark",
    country == "nl" ~ "netherlands",
    country %in% c("untied states", "united state", "united stated", "united sates", "unites states", "united statws", "united state of america") ~ "united states", # There must be a regular expression solution to this mess
    TRUE ~ country
    ), country_code = countrycode(country, origin = 'country.name', destination = 'iso3c')) # %>%
 # filter(is.na(country_code) == TRUE) # Testing only, shows how far to go to clean this data ...
```

```{r}
survey_clean %>%
  mutate(country_name = countrycode(country_code, origin = 'iso3c', destination = 'country.name')) %>%
  group_by(country_name) %>%
  summarise(count = n()) %>%
  top_n(n = 10, wt = count) %>%
  ggplot(aes(x = reorder(country_name, count), y = count)) + geom_col() + coord_flip()
```





# Step 1: Fix up categories for age, education
# Step 2: Normalise/standardise currencies, create total salary




## References
