---
title: "Tidy Tuesday 2021-21 Ask a manager survey"
output:
  github_document: default
  html_notebook: default
  html_document:
    df_print: paged
    self_contained: false
bibliography: ["../tidytuesday.bib"]
nocite: '@*'
---

## Introduction

My first serious attempt at using R to clean real data. 

## Setup
```{r setup, message = FALSE}
library(tidytuesdayR)
library(tidyverse)
library(countrycode)
library(scales)
```

<!-- ## Import data - first time -->
<!-- ```{r get_the_data, cache = TRUE, message = FALSE} -->
<!-- survey <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-18/survey.csv') -->
<!-- write_csv(survey, 'data/survey.csv') -->
<!-- rm tuesdata -->
<!-- ``` -->

## Import data - subsequent times
```{r get_the_data_local, cache = TRUE, message = FALSE}
survey <- read_csv('data/survey.csv')
survey_clean <- survey
glimpse(survey)
```

## Data cleaning - this may take some time ...

### Gender
Let's start by looking at the gender variable in our dataset.
```{r}
survey %>% 
  count(gender, sort = TRUE)
```
While this looks pretty clean, there are two immediate issues.
First, there is a single "prefer not to answer" observation even though there are numerous observations described as  "Other or prefer not to answer". Since there is no way to distinguish between survey respondents who are identifying with an "other" gender not listed and those that "prefer not to answer", this group of observations cannot be separated. To simplify, let's roll the single "prefer not to answer" observation into this broader group.
Second, these observations currently use the character data type, and it would be helpful for future analyses to have gender as a categorical variable.
Once tidied up, added a percentage column in case of future plots.

```{r clean_gender}
# Clean the gender data, by collapsing one category into another.
survey_clean <- survey %>%
  mutate(gender_clean = fct_collapse(gender, "Other or prefer not to answer" = "Prefer not to answer")) %>%
# Clean up the temporary columns so variable name is 'gender' and set data type to categorical.
  select(-gender) %>%
  mutate(gender = as.factor(gender_clean)) %>%
  select(-gender_clean)
  
# Calculate the percentages of each category, sort the data and plot it.
survey_clean %>%
  count(gender) %>%
  mutate (percent = round(n / sum(n) * 100, 1)) %>%
  arrange(desc(percent)) %>%
  ggplot(aes(x = reorder(gender, -percent), y = percent)) + geom_col()
```

#### Questions
1. I had a go at using stringr to clean this as strings, but it was pretty painful. This forcats package solution seemed easier and probably better suited to the categorical data that I was trying to wrangle. Are there positives/negatives to either approach ?
2. Is there merit in collapsing the "Other or prefer not to answer" group of observations into "NA". If so, how ..? I got stuck trying to mutate the same column multiple times and R wasn't happy.

### Country
Let's look at country now.

```{r country_view}
survey %>% 
  count(country, sort = TRUE)
```
What a mess. Let's have a go at tidying some of these up ...

```{r}

survey_clean <- survey_clean %>%
  mutate(country = str_to_lower(country)) %>%
  mutate(country = case_when(
    str_detect(country, "england") ~ "united kingdom",
    str_detect(country, "scotland") ~ "united kingdom",
    str_detect(country, "wales") ~ "united kingdom",
    str_detect(country, "northern ireland") ~ "united kingdom",
    country %in% c("englang", "united kindom", "unites kingdom", "u.k.", "gb") ~ "united kingdom",
    country %in% c("can", "canda", "csnada", "canad") ~ "canada",
    country == "danmark" ~ "denmark",
    country == "australi" ~ "australia",
    country == "nz" ~ "new zealand",
    country == "brasil" ~ "brazil",
    country %in% c("nl", "nederland") ~ "netherlands",
    country %in% c("untied states", "united state", "united stated", "united sates", "unites states", "united statws", "united state of america", "united statew", "united stares", "unitef stated", "uniyed states", "unted states", "usaa", "california", "united stateds", "unite states") ~ "united states",
    TRUE ~ country
    ), country_code = countrycode(country, origin = 'country.name', destination = 'iso3c'))

survey_clean %>%
  mutate(country_name = countrycode(country_code, origin = 'iso3c', destination = 'country.name')) %>%
  group_by(country_name) %>%
  summarise(count = n()) %>%
  top_n(n = 10, wt = count) %>%
  ggplot(aes(x = reorder(country_name, count), y = count)) + geom_col() + coord_flip()

```
## Finance data

```{r}
# Read in a list of exchange rates. In future I would want this to be dynamically generated, using an API like fixerapi, but for now, I'd like the exchange rates to stay the same, so it I can more easily see errors I'm creating.
rates <- read_csv("data/exchange.csv")

survey_clean %>%
  select(annual_salary, currency, currency_other) %>%
  filter(currency != "USD" & is.na(currency) == FALSE) %>%
  count(currency, sort = TRUE)
  # Returns non-USD entries

survey_clean %>%
  select(annual_salary, currency, currency_other) %>%
  filter(currency == "Other" & is.na(currency_other) == FALSE)
  # Returns entries where the currency was "Other", so hopefully there's something in the "Other" field to mutate across.

survey_clean %>%
  select(annual_salary, currency, currency_other) %>%
  filter(currency == "USD" & is.na(currency_other) == FALSE)
  # Returns USD entries where there was also something in the "Other" field.

survey_clean %>%
  select(annual_salary, currency, currency_other) %>%
  inner_join(rates, by = c("currency" = "source")) %>%
  filter(currency == "USD" & is.na(currency_other) == TRUE & annual_salary != 0) %>%
  mutate(usd_salary = annual_salary * rate) %>%
  select(usd_salary) %>%
  arrange(desc(usd_salary)) %>%
  ggplot(aes(usd_salary)) + geom_histogram(binwidth = 5000) + scale_x_continuous("Annual salary ($US)", label = label_dollar(), limits = c(1000, 500000))
# Where did the 2 missing values come from, given filtering for NA above?
# What are the 90 rows with 'non-finite values'? 

```
## Going to try something crazy ... fitting a linear model to the annual salary calculations, using age (initially)

```{r}

input <- survey_clean %>%
  select(how_old_are_you, annual_salary, currency, currency_other) %>%
  inner_join(rates, by = c("currency" = "source")) %>%
  filter(is.na(currency_other) == TRUE & annual_salary != 0 & how_old_are_you != "under 18") %>%
  mutate(usd_salary = annual_salary * rate) %>%
  filter(usd_salary < 1000000) %>%
  mutate(age = case_when(
    how_old_are_you == "18-24" ~ 21,
    how_old_are_you == "25-34" ~ 29.5,
    how_old_are_you == "35-44" ~ 39.5,
    how_old_are_you == "45-54" ~ 49.5,
    how_old_are_you == "55-64" ~ 60,
    how_old_are_you == "65 or over" ~ 65,
    )
  ) %>%
  mutate(age2 = age * age) %>%
  select(age, age2, usd_salary) %>%
  arrange(desc(usd_salary))

model_plot <- ggplot(data = input, aes(x = age, y = usd_salary)) + geom_point() + scale_y_continuous("Annual salary (US$)", label = label_dollar())

model <- lm(usd_salary ~ age + age2, data = input)

model_plot + geom_smooth(method = "lm", formula = y ~ poly(x, 2))

model_plot
```
```{r}
survey_clean %>%
  select(how_old_are_you, annual_salary, currency, currency_other) %>%
  inner_join(rates, by = c("currency" = "source")) %>%
  filter(is.na(currency_other) == TRUE & annual_salary != 0 & how_old_are_you != "under 18") %>%
  mutate(usd_salary = annual_salary * rate) %>%
  group_by(how_old_are_you) %>%
  summarize(avg_salary = mean(usd_salary)) %>%
  ggplot(aes(x = how_old_are_you, y = avg_salary)) + geom_col() + scale_y_continuous("Average annual salary (US$)", label = label_dollar()) + scale_x_binned("Age group")
```


## References
