---
title: "Tidy Tuesday: 2021-24 Commercial Fishing"
output:
  github_document: default
  html_notebook: default
  html_document:
    df_print: paged
    self_contained: false
bibliography: ["../tidytuesday.bib"]
nocite: '@*'
---

http://www.glfc.org/fishstocking/images/dbstruct.gif

```{r read in data}
library(tidyverse)
library(lubridate)
fishing_raw <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-08/fishing.csv')
stocked_raw <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-08/stocked.csv')
```

Appear to have lost 36320 rows in importing the stock data. Still have 56232 rows. Though `r round((36320 / (56232 + 36320) * 100), 0)`% loss seems like a lot.

On a quick inspection using problems() ... 

```{r}
problems(stocked_raw)
```
... it looks like:
* the tag_no variable is being loaded as a logical variable, when it should be a number or a series of numbers ...

```{r}
stocked_raw <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-08/stocked.csv', col_types = "ddddffccdddccccddcdccccdcdcccdc")
```
No problems ! 
Here's a list of the column types ...
SID, d
YEAR, d
MONTH, d
DAY, d
LAKE, f
STATE_PROV, f
SITE, c 
ST_SITE, c
LATITUDE, d
LONGITUDE, d
GRID, d
STAT_DIST, c
LS_MGMT, c
SPECIES, c
STRAIN, c
NO_STOCKED, d
YEAR_CLASS, d
STAGE, c
AGEMONTH, d
MARK, c
MARK_EFF, c
TAG_NO, c
TAG_RET, d to c
LENGTH, d
WEIGHT, d to c
CONDITION, d
LOT_CODE, c
STOCK_METH, c
AGENCY, c
VALIDATION, d
NOTES, c

## Create a date field (where possible)
```{r}
stocked_raw %>%
  mutate(DATE = make_date(YEAR, MONTH, DAY))
  
```
We're really only interested in:
- chinook salmon aka king salmon, which is the main target for fishing
- alewife or herring, the main food for chinook salmon

Something weird in how this data is presented. Lake == michigan but has US totals row ..?

```{r}
fishing_raw %>%
  filter(!is.na(values), 
      year > 1966,# year first stocked
      species %in% c("Chinook Salmon", "Coho Salmon", "Pacific Salmon", "Alewife") # various names used
      # geography/totals?
      ) %>%
  group_by(lake, year) %>%
  summarise(total_value = sum(values)) %>%
  ggplot(aes(x = year, y = total_value)) +
    geom_line() +
    facet_wrap(~ lake)
```
```{r}
fishing_raw %>%
  filter(!is.na(values)) %>%
  mutate(species = str_to_lower(species)) %>%
  mutate(species = case_when(
    species %in% c("amercian eel") ~ "american eel",
    species %in% c("bullhead", "bullheads", "channel catfish") ~ "channel catfish and bullheads",
    species %in% c("crappies", "crappie", "rock bass") ~ "rock bass and crappie",
    species %in% c("herring", "cisco", "chubs", "cisco and chub", "lake whitefish") ~ "cisco and chubs",
    species %in% c("drum", "sheepshead", "freshwater drum") ~ "freshwater drum or sheepshead",
    species %in% c("walleye", "blue pike") ~ "walleye and blue pike",
    TRUE ~ species
    )
  ) %>%
  group_by(lake, year, species) %>%
  summarise(total_value = sum(values)) %>%
  ggplot(aes(x = year, y = total_value)) +
    geom_line() +
    facet_wrap(~ species)
```

# Where fish have been stocked

```{r}

# Doesn't work ... weirdness with ggmap package and vectors

library(sf)
library(ggmap)

stocked_raw %>% 
  filter(!is.na(LONGITUDE) == TRUE & !is.na(LATITUDE) == TRUE) %>% 
  filter(LONGITUDE <= -50 & LATITUDE >= 40) %>% 
  st_as_sf(coords = c("LONGITUDE","LATITUDE"), crs = 4326) %>% 
  st_transform(., crs = 3857) -> stocked_geo

bbox <- st_bbox(stocked_geo)

base_layer <- get_stamenmap(
  bbox = c(
    left = bbox$xmin[[1]],
    bottom = bbox$ymin[[1]], 
    right = bbox$xmax[[1]], 
    top = bbox$ymax[[1]]
    ),
  maptype  = "terrain",
  zoom = 15
)

stocked_plot <- stocked_geo %>% 
  mutate(DATE = make_date(YEAR, MONTH, DAY)) %>%
  dplyr::select(DATE, LAKE, SPECIES, NO_STOCKED)

ggmap(base_layer) +
  #  geom_raster(aes(base_layer)) +
    geom_sf(data = stocked_plot, aes(size = NO_STOCKED, color = SPECIES))

```



# References